Parallel Multipoint Approximation Method for Large Scale Optimization Problems \thanks{This study was supported by the Russian Science Foundation, project No. 16-11-10150.}

### 2017-11-24 2018/PAVT_Toropov/PAVT2018_MAM.tex
Pred-final

    Denote the approximate objective function and constraints by $\widetilde{F}^k_0(x)$ and $\widetilde{F}^k_j(x)$, respectively.
    \item The original optimization problem (\ref{eq:problem}) is replaced by the following problem:
    \begin{equation}
      \label{eq:problem_approx}
      \begin{array}{c}
      \min\limits_{a_i^k \le x_i ^k\le b_i}\widetilde{F}^k_0(x) \\
      s.t.\; \widetilde{F}^k_j( x) \le 1,\; j=1,\dots ,M,
      \end{array}
    \end{equation}
    The approximate problem (\ref{eq:problem_approx}) is solved using Sequential Quadratic Programming (SQP) and the solution of this problem determines the center of the next trust region.
    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $ x^k$, and on the size of the current trust region \cite{KeulenToropovMarkine1996}.
    \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $ x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations). If the termination criterion is satisfied, the algorithm proceeds to step 3. Otherwise, it returns to step 2.
  \end{enumerate}
  \item Optimization terminates. The obtained approximation to the solution of the problem (\ref{eq:problem}) is $ x^{k+1}$.
\end{enumerate}

@@

The deflection $y_i$ at the right hand of the $i$-th segment is given by the following recursive formulas:
\begin{displaymath}
  \begin{array}{c}
    y_0=y_0'=0, \\
    y_i'=\frac{P\cdot l_i}{E\cdot I_i}\Big[ L+\frac{l_i}{2}-\sum\limits_{j=1}^i l_j\Big]+y_{i-1}', \\
    y_i=\frac{P\cdot l_i^2}{2E\cdot I_i}\Big[L-\sum\limits_{j=1}^i l_j + \frac{2l_i}{3}\Big]+y_{i-1}'l_i+y_{i-1}.
  \end{array}
\end{displaymath}
The moment of inertia of a segment is $I_i=\frac{b_i h_i^3}{12}$ and the bending moment at the left end is $M_i=P[L+l_i- \sum_{j=1}^{i}  l_j ]$. The maximum bending stress in the segment $i$ is then given by the following formula:
\begin{displaymath}
  \sigma_i=\frac{M_i h_i}{2I_i}
\end{displaymath}

We are looking for a design with smallest volume $V = \sum_{i=1}^N b_i h_i l_i$. The widths $b_i$ vary from 1.0 to 10.0 cm and the heights $h_i$ from 5.0 to 100.0 cm. The optimization problem is formulated as follows:
\begin{displaymath}
  \begin{array}{c}
    \min\limits_{ b,  h}V( b,  h) \\
    s.t.\;1.0\le b_i \le 10.0, \\
    5.0 \le h_i \le 100.0, \\
    y_N\le 2/5, \\
    \sigma_i \le \bar{\sigma}=14000, \\
    \frac{h_i}{b_i}\le 20.
  \end{array}
\end{displaymath}

@@

MAM obtained the solution $V = 63935.360\; cm^3$, using 1500 function evaluations (as compared to almost 8000 evaluations used by SQP). The number of points in a trust region used to build approximations was 150. The optimal values of design variables obtained by MAM are given below:
\[
\begin{matrix}
	b=&	\left[3.238,\right.&3.204,&3.202,&3.170,&3.176,&3.128,&3.108,&3.089,&3.057,&3.003 \\  
		& 3.027,&2.986,&2.959,&2.952,&2.885,&2.855,&2.864,&2.847,&2.803,&2.762 \\  
		& 2.737,&2.722,&2.630,&2.645,&2.590,&2.558,&2.547,&2.480,&2.693,&2.391 \\ 
		& 2.368,&2.310,&2.307,&2.227,&2.176,&2.149,&2.106,&2.016,&2.007,&1.925 \\  
		& 1.864,&1.843,&1.758,&1.635,&1.582,&1.934,&1.332,&1.173,&1.026,&\left.2.419\right]
\end{matrix}
\]
\[
\begin{matrix}
	h=& \left[64.764,\right.&64.083,&64.036,&63.392,&63.518,&62.566,&62.153,&61.772,&61.149,& 60.054  \\
		& 60.534,&59.728,&59.182,&59.033,&57.695,&57.105,&57.274,&56.943,&56.057,&55.233  \\
		& 54.747,&54.433,&52.590,&52.902,&51.802,&51.150,&50.943,&49.597,&51.567,&47.811  \\
		& 47.373,&46.190,&46.140,&44.543,&43.525,&42.991,&42.119,&40.308,&40.141,&38.506  \\
		& 37.268,&36.839,&35.153,&32.702,&31.615,&28.304,&26.642,&23.447,&20.536,& \left.9.417\right]
\end{matrix}
\]

@@

Next, let us compare the worktime of the sequential and parallel algorithms when solving the large scale problems. 
The dimensionality of the problem being solved $N$ was varied from 100 up to 1000 design variables that corresponds to 50 to 500 segments of the cantilevered beam. The number of points in a trust region used to build the approximations was $1.5 N$. 
Both algorithms was run on a single node of the cluster (the specifications of the node are listed below), the parallel algorithm employed all 16 processors cores available. Since the time of computing of the objective 
function and constraints in the test problem being solved was negligible, these ones were computed on the same node.

@@

\begin{table}
	\caption{Time and speedup}
	\label{tab:1}
	\center
	\begin{tabular}{cccc}
		\hline\noalign{\smallskip}
		$N$ & $T_{MAM}$ & $T_{PMAM}$ & Speedup \\
		\hline\noalign{\smallskip}
		100 & 15.4 & 11.9 & 1.3  \\
		200 & 167 &  89 &  1.9 \\
		500 & 5151 &  1520 & 3.4  \\
		1000 & 67674 &  13771 & 4.9  \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Function evaluations}
	\label{tab:2}
	\center
	\begin{tabular}{cccc}
		\hline\noalign{\smallskip}
		$N$ & $I_{MAM}$  & $F_{MAM}$ & $F_{SQP}$ \\
		\hline\noalign{\smallskip}
		100 & 10 &  2201 & 9901 \\
		200 & 10 &  4402 & 28143 \\
		500 &  &   &  115734 \\
		1000 &  &   & 305308 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

>>>>

    Denote the approximate objective function and constraints by $\widetilde{F}^k_0(x)$ and $\widetilde{F}^k_j(x)$, respectively.
    \item The original optimization problem (\ref{eq:problem}) is replaced by the following problem:
    \begin{equation}
      \label{eq:problem_approx}
      \begin{array}{c}
      \min\limits_{a_i^k \le x_i ^k\le b_i}\widetilde{F}^k_0(x) \\
      s.t.\; \widetilde{F}^k_j( x) \le 1,\; j=1,\dots ,M.
      \end{array}
    \end{equation}
    The approximate problem (\ref{eq:problem_approx}) is solved using Sequential Quadratic Programming (SQP) and the solution of this problem determines the center of the next trust region.
    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $ x^k$, and on the size of the current trust region \cite{KeulenToropovMarkine1996}.
    \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $ x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations). If the termination criterion is satisfied, the algorithm proceeds to step 3. Otherwise, it returns to step 2.
  \end{enumerate}
  \item Optimization terminates. The obtained approximation to the solution of the problem (\ref{eq:problem}) is $ x^{k+1}$.
\end{enumerate}

@@

The deflection $y_i$ at the right hand of the $i$-th segment is given by the following recursive formulas:
\begin{displaymath}
  \begin{array}{c}
    y_0=y_0'=0, \\
    y_i'=\frac{P\cdot l_i}{E\cdot I_i}\Big[ L+\frac{l_i}{2}-\sum\limits_{j=1}^i l_j\Big]+y_{i-1}', \\
    y_i=\frac{P\cdot l_i^2}{2E\cdot I_i}\Big[L-\sum\limits_{j=1}^i l_j + \frac{2l_i}{3}\Big]+y_{i-1}'l_i+y_{i-1}.
  \end{array}
\end{displaymath}
The moment of inertia of a segment is $I_i={b_i h_i^3}/{12}$ and the bending moment at the left end is $M_i=P[L+l_i- \sum_{j=1}^{i}  l_j ]$. The maximum bending stress in the segment $i$ is then given by the following formula:
\begin{displaymath}
  \sigma_i=\frac{M_i h_i}{2I_i}
\end{displaymath}

We are looking for a design with smallest volume $V = \sum_{i=1}^N b_i h_i l_i$. The widths $b_i$ vary from 1.0 to 10.0 cm and the heights $h_i$ from 5.0 to 100.0 cm. The optimization problem is formulated as follows:
\begin{displaymath}
  \begin{array}{c}
    \min\limits_{ b,  h}V( b,  h) \\
    s.t.\;1.0\le b_i \le 10.0, \\
    5.0 \le h_i \le 100.0, \\
    y_N\le 2.5, \\
    \sigma_i \le \bar{\sigma}=14000, \\
    \frac{h_i}{b_i}\le 20.
  \end{array}
\end{displaymath}

@@

MAM obtained the solution $V = 63935.360\; cm^3$, using 2201 function evaluations (as compared to almost 10000 evaluations used by SQP). The number of points in a trust region used to build approximations was 200. The optimal values of design variables obtained by MAM are given below:
\[
\begin{matrix}
	b=&	\left[3.238,\right.&3.204,&3.202,&3.170,&3.176,&3.128,&3.108,&3.089,&3.057,&3.003 \\  
		& 3.027,&2.986,&2.959,&2.952,&2.885,&2.855,&2.864,&2.847,&2.803,&2.762 \\  
		& 2.737,&2.722,&2.630,&2.645,&2.590,&2.558,&2.547,&2.480,&2.693,&2.391 \\ 
		& 2.368,&2.310,&2.307,&2.227,&2.176,&2.149,&2.106,&2.016,&2.007,&1.925 \\  
		& 1.864,&1.843,&1.758,&1.635,&1.582,&1.934,&1.332,&1.173,&1.026,&\left.2.419\right]
\end{matrix}
\]
\[
\begin{matrix}
	h=& \left[64.764,\right.&64.083,&64.036,&63.392,&63.518,&62.566,&62.153,&61.772,&61.149,& 60.054  \\
		& 60.534,&59.728,&59.182,&59.033,&57.695,&57.105,&57.274,&56.943,&56.057,&55.233  \\
		& 54.747,&54.433,&52.590,&52.902,&51.802,&51.150,&50.943,&49.597,&51.567,&47.811  \\
		& 47.373,&46.190,&46.140,&44.543,&43.525,&42.991,&42.119,&40.308,&40.141,&38.506  \\
		& 37.268,&36.839,&35.153,&32.702,&31.615,&28.304,&26.642,&23.447,&20.536,& \left.9.417\right]
\end{matrix}
\]

@@

Next, let us compare the worktime of the sequential and parallel algorithms when solving the large scale problems. 
The dimensionality of the problem being solved $N$ was varied from 100 up to 1000 design variables that corresponds to 50 to 500 segments of the cantilevered beam. The number of points in a trust region used to build the approximations was $2 N$. 
Both algorithms was run on a single node of the cluster (the specifications of the node are listed below), the parallel algorithm employed all 16 processors cores available. Since the time of computing of the objective 
function and constraints in the test problem being solved was negligible, these ones were computed on the same node.

\begin{table}
	\caption{Time and speedup}
	\label{tab:1}
	\center
	\begin{tabular}{cccc}
		\hline\noalign{\smallskip}
		$N$ & $T_{MAM}$ & $T_{PMAM}$ & Speedup \\
		\hline\noalign{\smallskip}
		100 & 15.4 & 11.9 & 1.3  \\
		200 & 167 &  89 &  1.9 \\
		400 & 2476 &  837 &  3.0 \\
%		500 & 5151 &  1520 & 3.4  \\
		600 & 9678 &  2777 &  3.5 \\
		800 & 27826 &  6948 &  4.0 \\
		1000 & 67674 &  13771 & 4.9  \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Function evaluations}
	\label{tab:2}
	\center
	\begin{tabular}{cccc}
		\hline\noalign{\smallskip}
		$N$ & $I_{MAM}$  & $F_{MAM}$ & $F_{SQP}$ \\
		\hline\noalign{\smallskip}
		100 & 10 &  2201 & 9901 \\
		200 & 10 &  4402 & 28143 \\
		400 & 11 &  9599 &  84211 \\
%		500 &  &   &  115734 \\
		600 & 10 &  13200 &  146046 \\
		800 & 10 &  17600 &  221079 \\
		1000 & 10 &  22000 & 305308 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

@@
